---
draft: true
title: 'Oceanographic Wave Measuring '
author: Aaron Willcox
date: '2019-08-18'
slug: oceanographic-wave-measuring
categories:
  - R
tags:
  - R Markdown
subtitle: ''
summary: ''
authors: []
lastmod: '2019-08-18T13:49:58+10:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

This is a simple graphical display of a dataset with wave data from Queensland Australia. 

```{r Librarys, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(hrbrthemes)
```

This dataset contains Measured/Calculated wave parameters. Measured and derived wave data from data collected by oceanographic wave measuring buoys anchored at Mooloolaba. Coverage period: 30 months.

Acknowledgements
This data comes from Queensland Government Data - [https://data.qld.gov.au/dataset](Link)

Date/TimeDate

+ Hs Significant wave height, an average of the highest third of the waves in a record
+ Hmax The maximum wave height in the record
+ Tz The zero upcrossing wave period
+ Tp The peak energy wave period
+ Peak Direction Direction (related to true north) from which the peak period waves are coming from
+ SST Approximation of sea surface temperature


```{r message=FALSE, warning=FALSE}

waves_df <-  read_csv("data/Coastal Data System - Waves (Mooloolaba) 01-2017 to 06 - 2019.csv")

# The date format is mixed so some are mdy and other dmy
# Which we will separate out below 
waves_df$`Date/Time` <- parse_date_time(waves_df$`Date/Time`, c("dmyHMS", "mdyHMS"), truncated = 3) # Change the date format


waves_df <- separate(waves_df, 'Date/Time', into = c("year", "month", "day"), sep = "-") # Separate out the year month and day


waves_df <- separate(waves_df, 'day', into = c("day", "time"), sep = " ") # separate out the day and time

head(waves_df)
```

#### Average Wave Height for 2017 to 2019 for each month

```{r message=FALSE, warning=FALSE}

waves_df %>% 
  group_by(year, month) %>% 
  summarise(Ave_wave_height = mean(Hs)) %>% 
  ggplot(aes(x = month, y = Ave_wave_height)) +
  geom_point(size = 2.2) +
  geom_line(aes(x = month, y = Ave_wave_height)) +
  facet_wrap(~year) +
  theme_bw()+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


waves_df %>% 
  group_by(year, month) %>% 
  summarise(Mean = mean(Hs)) %>% 
ggplot(aes(x = month, y = Mean, group = year)) + 
  geom_line(aes(color = year), size = 1) +
  theme_ft_rc() +
  xlab("Month") + 
  ylab("Mean Wave Height") +
  ggtitle("Average Wave Height 2017 to 2019")

```

Correlations between variables

```{r}

waves_cor <- waves_df %>% select(5:10)

waves_cor <- cor(waves_cor)

corrplot::corrplot(waves_cor, is.corr = FALSE)

```

