---
title: Global Suicides 1995 to 2015
author: Aaron Willcox
date: '2019-06-01'
slug: global-suicides-1995-to-2015
categories:
  - R
tags:
  - R Markdown
subtitle: ''
summary: ''
authors: []
lastmod: '2019-06-01T20:03:33+02:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
fig_width: 12
---

## R Markdown

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
library(modelr)
library(broom)
library(scales)
library(cowplot)
```

#### Import dataset

```{r Data Import, message=FALSE, warning=FALSE}
suicide_raw <- read_csv("data/master.csv")
suicide_df <- suicide_raw %>% 
  select(country
         , year
         , "No_of_Suicides" =`suicides_no`
         , "ave_suicide_100k" = `suicides/100k pop`
         , sex
         , age
         , population
         , "HDI" = `HDI for year`
         , "GDP_per_year" = `gdp_for_year ($)`
         , "GDP_per_capita" = `gdp_per_capita ($)`
         , generation)

# converts to factors
suicide_df$sex <- factor(suicide_df$sex, levels = c("male", "female"))
suicide_df$year <- factor(suicide_df$year)
suicide_df$generation <- factor(suicide_df$generation)
suicide_df$country <- factor(suicide_df$country)

suicide_df$generation <- factor(suicide_df$generation, 
                   ordered = T, 
                   levels = c("G.I. Generation", 
                              "Silent",
                              "Boomers", 
                              "Generation X", 
                              "Millenials", 
                              "Generation Z"))

```

#### Explore dataset 

Take a look at the top 50 global number of suicides from 1985 to 2015 and split by sex. 


### Total Amount of Suicides from 1995 to 2015

```{r, fig.width=10, fig.height=10}


suicide_df %>% 
  group_by(country, sex) %>% 
  summarise(Suicide_total = sum(No_of_Suicides)) %>% 
  arrange(desc(Suicide_total))  %>% 
  head(50) %>%  
  ggplot(aes(x = reorder(country, -Suicide_total), y = Suicide_total)
  ) +
  geom_col(show.legend = FALSE, alpha = 0.7) +
  coord_flip() +
  facet_wrap(~sex) +
  labs(x = "Country", y = "Suicides 1995 to 2015", title = "Total Number of Suicides from 1985 to 2015") 

```

The Russian Federation has had the most amount of people commit suicide between 1985 to 2015. 
Following this, United States and Japan have a high number of suicides with Males commiting suicide
more than Females.  

#### Average suicide rate per population of 100,000

```{r, fig.width=10, fig.height=12}


ave_suicide_100k_country <- suicide_df %>% 
  group_by(country) %>% 
  summarise(n = n(), Ave = mean(ave_suicide_100k)) %>% 
  arrange(desc(Ave))


ave_suicide_100k_year <- suicide_df %>% 
  group_by(country, year) %>% 
  summarise(n = n(), Ave = mean(ave_suicide_100k)) %>% 
  arrange(desc(Ave))


# reorder the country 
ave_suicide_100k_country$country <- factor(ave_suicide_100k_country$country, 
                          ordered = T, 
                          levels = rev(ave_suicide_100k_country$country))

average_suicide_no <- (mean(ave_suicide_100k_country$Ave))

ave_suicide_100k_country %>% 
  ggplot(aes(x = country, y = Ave)) + 
  geom_bar(stat = "identity") + 
  geom_hline(yintercept = average_suicide_no, linetype = 2, color = "grey35", size = 1) +
  labs(title = "Global suicides per 100k, by Country",
       caption = "1985 to 2015",
       x = "Country", 
       y = "Suicides per 100k") +
  coord_flip() +
  scale_y_continuous(breaks = seq(0, 45, 2)) 
```


#### Total Average Suicide Rate per Generation between 1985 to 2015


  * Gen Z, iGen, or Centennials: Born 1996 – TBD
  * Millennials or Gen Y: Born 1977 – 1995
  * Generation X: Born 1965 – 1976: 15 - 24 years
  * Baby Boomers: Born 1946 – 1964
  * Traditionalists or Silent Generation: Born 1945 and before


```{r}
ave_suicide_100k_gen <- suicide_df %>% 
  group_by(generation) %>% 
  summarise(n = n(), Ave = sum(No_of_Suicides)) %>% 
  arrange(desc(Ave))

ave_suicide_100k_gen$generation <- factor(ave_suicide_100k_gen$generation, 
                          ordered = T, 
                          levels = rev(ave_suicide_100k_gen$generation))

ggplot(data = ave_suicide_100k_gen) +
  aes(x = generation, weight = Ave) +
  geom_bar(fill = "#0c4c8a") +
  labs(title = "Total Suicides per Generation",
    x = "Generation",
    y = "Average Suicide Rate",
    subtitle = "1985 to 2015") +
  theme_bw() +
  coord_flip()

```

#### Total Amount of Suicides per country by Age (1985 to 2015)

```{r fig.height=12, fig.width=12, message=FALSE, warning=FALSE}

x <- suicide_df %>% 
  group_by(country, age) %>% 
  summarise(total = sum(No_of_Suicides)) %>% 
  filter(total > 2000) %>% 
  arrange(desc(total))

ggplot(x, aes(x = reorder(country, - total), y = total, fill = age)) + 
  geom_bar(stat = "identity") + 
  labs(title = "Total Amount of Suicides per country by Age (1985 to 2015)",
       subtitle = "Filtered by Suicides greater than 2000",
       x = "Country", 
       y = "Total Amount of Suicides", 
       fill = "age") +
  coord_flip() + 
  theme(legend.position = "bottom")
```

Age group of 15 to 24 years has been a consistant marker for suicides. To Explore this further, we can look at 
a scatter plot ages over time. Will also remove the 5 to 14 years age group. 

```{r, fig.width=10, fig.height=8}

# line graph by age group



all_age_group <- suicide_df %>%
  group_by(year, age) %>% 
 filter(!(age %in% "5-14 years")) %>% 
  filter(!(year %in% "2016")) %>% 
  #na.omit() %>% 
  summarise(Num = sum(No_of_Suicides))

ggplot(all_age_group) +
 aes(x = year, y = Num, fill = age, colour = age, group = age) +
 geom_point(size = 2.2) +
  geom_line() +
 labs(x = "Year", y = "Total Num Suicides", title = "Total Suicides per Age Group") +
 theme_bw()+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```

#### Regression model 

We can use a regression model to see which age group is most likely to attempt suicide.  

```{r}

model <- suicide_df %>% 
  lm(No_of_Suicides ~ GDP_per_year + age + sex + generation, data = .) 

model %>% 
  tidy(conf.int = TRUE) %>% 
  filter(term != "(Intercept)") %>%
  mutate(term = fct_reorder(term, estimate)) %>% # almost always remove this
  ggplot(aes(estimate, term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high))

# model %>% 
#  augment(data = suicide_df) %>% 
#  ggplot(aes(.fitted, No_of_Suicides)) +
#  geom_point(alpha = .1)

tidy(anova(model)) %>% 
  mutate(Unique = sumsq / sum(sumsq))

```

A rough look at the effects supports previous assumptions that being male and aged between 35 - 54 years are at a higher risk of suicide. Although there might well be some collinearity between generation and age groups as these are essentially capturing similar items. I will also remove the 5 -14 year old age group as this doesnt seem to be adding anything to the model. 

```{r message=FALSE, warning=FALSE}



model2 <- suicide_df %>% 
  filter(!(age %in% "5-14 years")) %>% 
  lm(No_of_Suicides ~ GDP_per_year + age + sex, data = .) 

model2 %>% 
  tidy(conf.int = TRUE) %>% 
  filter(term != "(Intercept)") %>%
  mutate(term = fct_reorder(term, estimate)) %>% # almost always remove this
  ggplot(aes(estimate, term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high))

tidy(anova(model2)) %>% 
  mutate(Unique = sumsq / sum(sumsq))
```

Interestly, men aged between 35 - 54 years old have the strongest effect on the number of suicides followed by the 55 to 74 year old age group then young adults. 

#### Take a look at Suicide Rate in Australia


```{r, fig.width=10, fig.height=12}

au_model <- suicide_df %>% 
  filter(country %in% "Australia") %>% 
  group_by(year, sex, age) %>% 
  summarise(No_of_Suicides = sum(No_of_Suicides), GDP_per_year = mean(GDP_per_year))

model2 <- au_model %>% 
  filter(!(age %in% "5-14 years")) %>% 
  lm(No_of_Suicides ~ GDP_per_year + age + sex, data = .) 

model2 %>% 
  tidy(conf.int = TRUE) %>% 
  filter(term != "(Intercept)") %>%
  mutate(term = fct_reorder(term, estimate)) %>% # almost always remove this
  ggplot(aes(estimate, term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high))

tidy(anova(model2)) %>% 
  mutate(Unique = sumsq / sum(sumsq))

```

Similar pattern in Australia. Men aged 35 - 54 years old have the highest suicide rate.

```{r message=FALSE, warning=FALSE}

age_group_35_54 <- suicide_df %>% 
  group_by(country, year, age) %>% 
 filter(age %in% "35-54 years") %>% 
  summarise(Num = sum(No_of_Suicides))


age_group_35_54 %>%  
  filter(country %in% "Australia") %>% 
ggplot() +
 aes(x = year, y = Num) +
 geom_point(size = 2, colour = "#0c4c8a") +
  geom_line() +
  labs(title = "Total Suicides in Australia 1985 to 2015", 
         subtitle = "Age Group 35 to 54 Years Old",
       x = "Year",
       y = "Number of Suicides") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```