---
title: OSL Workshop Solomons Paradox
author: Aaron Willcox
date: '2019-06-01'
slug: osl-workshop-solomons-paradox
categories:
  - R
tags:
  - Academic
  - R Markdown
  - Workshop
subtitle: ''
summary: ''
authors: []
lastmod: '2019-06-01T20:20:58+02:00'
featured: yes
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

## Does reasoning about personal problems improve with psychological distance?

#### STUDY DESCRIPTION

Solomon’s paradox describes the tendency for people to reason more wisely about other people’s
problems compared to their own. One potential explanation for this paradox is that people tend to view
other people’s problems from a more psychologically distant perspective, whereas they view their own
problems from a psychologically _immersed perspective_ . For example, imagine two friends, Beth and Zoe,
are discussing Zoe’s relationship problems. Beth’s distance allows her to see that Zoe’s relationship is
doomed, so she can offer her friend sage advice for how to proceed with her relationship. Zoe’s
immersion in her own relationship may lead her to have a hard time reasoning wisely, because she may be
worried that she will need to find a new apartment if she breaks up with her boyfriend.

What if, however, Zoe was able to take a more psychologically _distanced perspective_ when contemplating
her relationship problems? Would she exhibit a _higher level of wisdom_, similar to what Beth was able to
show? To test this possibility, Grossmann and Kross (2014) asked romantically-involved participants to
_think about a situation_ in which their partner cheated on them ( __self condition__ ) or a friend’s partner
cheated on their friend ( __other condition__ ). Participants were also instructed to take a first-person
perspective ( __immersed condition__ ) by using pronouns such as I and me, or a third-person perspective
( ___distanced condition___ ) by using pronouns such as he and her.

Participants were __120__ university students who were involved in monogamous, heterosexual romantic
relationships, and participants were randomly assigned to condition. After contemplating the infidelity
scenario described above with the assigned perspective, participants responded to various questions
designed to assess __wise reasoning__.


Grossmann, I., & Kross, E. (2014). Exploring Solomon’s paradox: Self-distancing eliminates self-other asymmetry in wise reasoning about close relationships in younger and older adults. Psychological Science, 25, 1571-1580.

> Given the context of this study, a dimension that could have been explored more was that of how wisdom differs
> to that of intelligence. In that inteliigence is more applied and wisdom is the ability to learn from an experience.
 

#### ANAYSIS

1. Open the data file (called Grossmann and Kross 2014 Study 2). Explore the data file. Note, you
will not analyze all of these variables. Try to find the variables that are relevant to the study
description above.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(car)
library(dplyr)
library(psych)
library(ggplot2)
library(knitr)
library(pander)

Psych_dist_df <- read.csv("data/Grossman and Kross 2014 Study 2.csv")
```


2. Conduct a one-way ANOVA to determine if there is a significant difference between the
conditions on wisdom.

Conditions: 
DV: Wise Reasoning


```{r message=FALSE, warning=FALSE}
# Use the names() function to print out the list of variable names
#names(Psych_dist_df)
# We are interested in the conditions and wisdom variables
# so we can use dplyr to select these and print a table to look at
# Can see some NA cases
#complete.cases(Psych_dist_df)

df <- Psych_dist_df[complete.cases(Psych_dist_df), ] # remove cases with NA and subset to new dataframe
df <- select(df, CONDITION, WISDOM) # filter out other variables and only use ones of interest

```


#### Descriptives


```{r message=FALSE, warning=FALSE}
Descriptives <- select(df, CONDITION, WISDOM) %>% 
  group_by(CONDITION) %>% 
  summarise(count = n()
            ,mean = mean(WISDOM)
            ,sd = sd(WISDOM))

# viewing the dataframe shows that the CONDITION is currently treated as an INT so we will 
# convert this to a factor with label names for clarity
df$CONDITION <- factor(df$CONDITION, labels = c("Self_immersed", "Self_dist", "other_immersed", "other_dist") ) # convert the CONDITION int to a factor


ggplot(data = df) +
  aes(x = CONDITION, y = WISDOM) +
  geom_boxplot(fill = "#26828e") +
  labs(title = "Solomon’s paradox",
    x = "Condition",
    y = "Wisdom",
    subtitle = "Subtitle") +
  theme_bw()



kable(Descriptives) %>% kableExtra::kable_styling()

```

#### Visual Analysis


#### Run ANOVA


```{r}

anova_model <- lm(WISDOM ~ CONDITION, data = df) #run an anova or lm
```

#### Run the model through a Anova function

```{r}
pander(Anova(anova_model, type="III")) # Ue the car package to run the anova
```


#### Levene's Test

```{r}
pander(leveneTest(anova_model)) # tests of homogeniety
```

#### Using the lm() function from earlier, printing the summary results in contrasts.

```{r}
pander(summary(anova_model)) # summary.lm gives us each level of the condition


```


#### Box plot
```{r}

ggplot(data = df 
       , aes(x = CONDITION, y = WISDOM, colour = CONDITION)) + 
  geom_boxplot(outlier.colour="red" # the geom_ must be on the same line
               , outlier.shape=16
               , outlier.size=2
               , notch=TRUE)  

```


3. Next, you want to determine whether the _self-immersed_ condition was significantly lower in
wisdom, relative to the _other-immersed_ and _other-distanced_ condition. Conduct a planned
contrast to test the typical Solomon’s paradox effect.

```{r, planned contrasts}
# planned contrasts
c1 <- c(2, 0, -1, -1)
planned_contrasts <-c1
contrasts(df$CONDITION) <- planned_contrasts
model1 <- aov(WISDOM ~ CONDITION, data = df)
Anova(model1, type = "III")
summary.aov(model1, split = list(CONDITION = list("Self_immersed" = 1, "Self_dist" = 2, "other_immersed" = 3, "other_dist" = 4)))
```


4. Now, you want to show that taking a distant perspective increases wisdom relative to taking an
immersed perspective when dealing with one’s own problems. Conduct a planned contrast to
determine whether _self-distancing_ results in significantly higher levels of wisdom, relative to
_self-immersion_.

```{r}
c2 <- c(1, -1, 0, 0) # self distance vs self immursion
planned_contrasts <-c2
contrasts(df$CONDITION) <- planned_contrasts
model2 <- aov(WISDOM ~ CONDITION, data = df)
Anova(model2, type = "III")
summary.aov(model2, split = list(CONDITION = list("Self_immersed" = 1, "Self_dist" = 2, "other_immersed" = 3, "other_dist" = 4)))
```


5. You also want to determine whether distancing vs. immersion increases wisdom when
contemplating other people’s problems. Conduct a planned contrast to compare the
_other-distance_ vs. _other-immersed_ conditions.

```{r}
c3 <- c(0, 0, 1, -1) # other immersed vs other distance
planned_contrasts <-c3
contrasts(df$CONDITION) <- planned_contrasts
model3 <- aov(WISDOM ~ CONDITION, data = df)
Anova(model3, type = "III")
summary.aov(model3, split = list(CONDITION = list("Self_immersed" = 1, "Self_dist" = 2, "other_immersed" = 3, "other_dist" = 4)))
```

6. Finally, you want to test whether self-distancing eliminates the increased wisdom typically found
in reasoning about others. Conduct a planned comparison to determine whether the
_self-distanced_ condition is significantly different from the _other-immersed_ and _other-distanced_
conditions.
```{r}
c4 <- c(0, 2, -1, -1) # self dist vs other immersed & other dist
planned_contrasts <-c4
contrasts(df$CONDITION) <- planned_contrasts
model4 <- aov(WISDOM ~ CONDITION, data = df)
Anova(model4, type = "III")
summary.aov(model4, split = list(CONDITION = list("Self_immersed" = 1, "Self_dist" = 2, "other_immersed" = 3, "other_dist" = 4)))
```


7. Prepare an APA-style results section to describe each of the analyses conducted above.


8. Generate a bar graph to depict the results for the one-way ANOVA. Don’t forget to include error
bars that reflect the +/- 1 standard error of the mean.

```{r}
ggplot(df, aes(CONDITION, WISDOM))+
  stat_boxplot( aes(CONDITION, WISDOM), 
    geom='errorbar', linetype=1, width=0.5)+  #whiskers
  geom_boxplot( aes(CONDITION, WISDOM),outlier.shape=1) +    
  stat_summary(fun.y=mean, geom="point", size=2) + 
  stat_summary(fun.data = mean_se, geom = "errorbar")
```

