---
title: Emotional Impact on Memory Part 2
author: Aaron Willcox
date: '2019-08-16'
slug: emotional-impact-on-memory-part-2
categories:
  - R
tags:
  - Emotion
  - R Markdown
  - research
subtitle: ''
summary: ''
authors: []
lastmod: '2019-08-16T10:30:23+10:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

This is the second part of the expriment on Pattern Separation. This processes the participants hits and misses into the induced negative or neutral group and displays some results. 

```{r Library, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggplot2)
#library(broom)
library(cowplot)
library(kableExtra)
library(DT)
library(readr)
```

#### Import

Raw CSV files with all IAPS database and images to be used in each task imported into an object. 

```{r message=FALSE, warning=FALSE}

raw_df <- read_csv("data/ret_task.csv"
                   , col_types = cols(CR = col_factor(levels = c("same",  "different"))
                   , Emotion = col_factor(levels = c("negative", "neutral"))
                  , Sex = col_factor(levels = c("male", "female"))))

raw_df$CorrectAnswer <- recode(raw_df$CorrectAnswer, a = "l", l = "a")
raw_df$Stimulus1.RESP <- recode(raw_df$Stimulus1.RESP, a = "l", l = "a")

# import tables 

RET_SEEN <- read_csv("data/RET_SEEN.csv")
RET_UNSEEN_NEW <- read_csv("data/RET_UNSEEN_NEW.csv")
RET_PS_a <- read_csv("data/RET_PS_a.csv")

```

#### Data Setup


+ 75 Negative Images

+ 75 Neutral Images


General d-prime score: p(hits|75) and then p(FA|75)

|         	| SAME(A)  	| DIFFERENT(L) 	|
|---------	|----------	|-------------	|
| NEW     	| MISS(FA) 	| HIT(CR)     	|
| OLD     	| MISS     	| MISS        	|
| SIMILAR 	| HIT      	| MISS        	|


##### Compute New Variables

Pattern separation scores were calculated by comparing the __CorrectAnswer__ response to the actual __Stimulus1.RESP__ and then allocated either a __hit__ or __miss__ where appropriate. Futher variables were then computed as a binary response  of __1__ for a hit or __0__ for a miss. 

Negative and Neutral conditions were also computed as when a participant responded to a hit when a image is in the negative condition. Neutral was the compliment of this. 


```{r message=FALSE, warning=FALSE}

raw_df <- raw_df %>% 
  select_all() %>% 
  mutate(
    Pattern_Sep = case_when(
      CorrectAnswer == Stimulus1.RESP & CorrectAnswer == "a"  ~ "Hit", 
      # Condition checks to match correct answer to 
      CorrectAnswer == Stimulus1.RESP & CorrectAnswer == "l" ~ "Hit",  
      # the participants response. Labelled as "hits"
      CorrectAnswer != Stimulus1.RESP & CorrectAnswer == "l" ~ "Miss", 
      # and "miss" for easy identifying. 
      CorrectAnswer != Stimulus1.RESP & CorrectAnswer == "a" ~ "Miss",
      TRUE  ~ "")) %>% 
  mutate(
    Score = case_when(
      Pattern_Sep == "Hit" ~ 1, # A column of hits and misses coded 1 for hit and 0 for miss
      Pattern_Sep == "Miss" ~ 0
    )
  )  %>% 
  mutate(
    Total_Hits = factor(case_when( # total counts for hits and misses
      Score == 1 ~ 1,
      Score == 0 ~ 0
    )),
    Total_Miss = factor(case_when(
      Score == 0 ~ 1,
      Score == 1 ~ 0
    ))
  )  %>% 
  mutate(Hit_similar = factor(ifelse(CorrectAnswer %in% c("a") &   
                                       # These columns count the appropriate hits and misses
                                Stimulus1.RESP == "a", 1, 0)),      
         # for each emotional condition. 
         Negative_Hit = factor(ifelse(Emotion %in% c("negative") &
                            Pattern_Sep == "Hit", 1, 0)),
         Negative_Miss = factor(ifelse(Emotion %in% c("negative") &
                              Pattern_Sep == "Miss", 1, 0)),
         Neutral_Hit = factor(ifelse(Emotion %in% c("neutral") &
                             Pattern_Sep == "Hit", 1, 0)),
         Neutral_Miss = factor(ifelse(Emotion %in% c("neutral") &
                                 Pattern_Sep == "Miss", 1, 0)))
 
raw_df$Picture <-  str_remove(raw_df$Picture, ".jpg") #remove the jpg extention


# check the column picture number and compare to the originally coded
# tasks use to setup the experiment. 
raw_df <- raw_df %>% 
  mutate(
    coding = case_when(
      Picture %in% RET_PS_a$V1 ~ "SIMILAR", 
      Picture %in% RET_UNSEEN_NEW$nr ~ "NEW", 
      Picture %in% RET_SEEN$nr ~ "OLD",
      TRUE ~ ""))

# Added pattern separation counts similar to Stark(2013)
raw_df <- raw_df %>% 
  mutate(pshits = ifelse(coding %in% c("SIMILAR") & Stimulus1.RESP == "a", 1, 0),
         psmiss = ifelse(coding %in% c("SIMILAR") & Stimulus1.RESP == "l", 1, 0),
         pscr = ifelse(coding %in% c("NEW") & Stimulus1.RESP == "l", 1, 0),
         psfa = ifelse(coding %in% c("NEW") & Stimulus1.RESP == "a", 1, 0))



```



#### Grouping Variables and Joining To Long Format: 

Temporary tables are then created per subject based on hits and misses. 

```{r Grouping Variables, message=FALSE, warning=FALSE}

a <- raw_df %>% group_by(Subject, Neutral_Hit) %>% 
  tally() %>% 
  filter(Neutral_Hit == 1) %>% 
  select(n) %>% 
  rename(Neutral_hits = n) 

b <- raw_df %>% group_by(Subject, Neutral_Miss) %>% 
  tally() %>% 
  filter(Neutral_Miss == 1) %>% 
  select(n) %>% 
  rename(Neutral_miss = n) 

c <- raw_df %>% group_by(Subject, Negative_Hit) %>% 
  tally() %>% 
  filter(Negative_Hit == 1) %>% 
  select(n) %>% 
  rename(Negative_hits = n) 

d <- raw_df %>% group_by(Subject, Negative_Miss) %>% 
  tally() %>% 
  filter(Negative_Miss == 1) %>% 
  select(n) %>% 
  rename(Negative_miss = n) 

e <- full_join(a, b)

f <- full_join(c, d)

g <- full_join(e, f)


h <- raw_df %>% 
  group_by(Subject) %>% 
  filter(Total_Hits == 1) %>% 
  tally() %>% 
  rename("Total_hits" = n)

i <- raw_df %>% 
  group_by(Subject) %>% 
  filter(Total_Miss == 1) %>% 
  tally() %>% 
  rename("Total_miss" = n)

j <- full_join(h, i)
k <- full_join(g, j)

l <- raw_df %>% 
  select(Subject, Sex, Stimulus1.RT) %>% 
  group_by(Subject, Sex) %>% 
  summarise(MeanRT = mean(Stimulus1.RT), )

Ret_task_a <- full_join(l, k)


m <- raw_df %>% 
  select(Subject, Sex, Emotion, Stimulus1.RT) %>% 
  filter(Emotion == "negative") %>% 
  group_by(Subject, Sex) %>% 
  summarise(Neg_RT = mean(Stimulus1.RT))

n <- raw_df %>% 
  select(Subject, Sex, Emotion, Stimulus1.RT) %>% 
  filter(Emotion == "neutral") %>% 
  group_by(Subject, Sex) %>% 
  summarise(Neut_RT = mean(Stimulus1.RT))

# ---
d1 <- raw_df %>% 
  group_by(Subject) %>% 
  filter(Emotion == "negative" & Hit_similar == "1") %>%
  tally() %>%
  rename("Neg_Sim_Hit" = n)

d2 <- raw_df %>% 
  group_by(Subject) %>% 
  filter(Emotion == "neutral" & Hit_similar == "1") %>%
  tally() %>%
  rename("Neut_Sim_Hit" = n)

Ret_task_c <- full_join(d1, d2)

# ---

Ret_task_b <- full_join(m, n)  

Ret_task_d <- full_join(Ret_task_b, Ret_task_c)

Ret_task_df_temp <- full_join(Ret_task_a, Ret_task_d) # temp join of data tables

pattern_sep_scores <- raw_df %>% 
  group_by(Subject, Sex) %>% 
  summarise(n_hits = sum(pshits)
            , n_miss = sum(psmiss)
            , n_cr = sum(pscr)
            , n_fa = sum(psfa)) %>% 
  mutate(n_hits = replace(n_hits, which(is.na(n_hits)), 0)) %>% 
  mutate(n_miss = replace(n_miss, which(is.na(n_miss)), 0)) %>% 
  mutate(n_cr = replace(n_cr, which(is.na(n_cr)), 0)) %>% 
  mutate(n_fa = replace(n_fa, which(is.na(n_fa)), 0))

 
 
pattern_sep_scores <- pattern_sep_scores  %>% 
  mutate(Perc_correct = (n_hits + n_cr)/(n_hits + n_miss + n_fa +n_cr)
         , Dprime = (n_hits/(n_hits + n_miss)-(n_fa/(n_fa + n_cr))))

Ret_task_df_join <-  full_join(Ret_task_df_temp, pattern_sep_scores)


datatable(Ret_task_df_join)
```



#### Validation Task Data

Validation database imported into an object

```{r message=FALSE, warning=FALSE}

val_task <- read_csv(file = "data/val_task.csv")

```

A table is created by grouping by each subject and then a new variable is computed for each emotion score dimension.
There is also response times for each image but wont be using those for this analysis. 

```{r message=FALSE, warning=FALSE}



val_df <- val_task %>% 
  group_by(Subject) %>% 
  summarise(Arousal = mean(arousal.RESP), Valence = mean(valence.RESP), Dominance = mean(dominance.RESP))

```


#### Full join of both tasks to one dataframe in wide format

```{r message=FALSE, warning=FALSE}


val_df$Subject <- as.factor(val_df$Subject)

Ret_task_df_join$Subject <- as.factor(Ret_task_df_join$Subject)

group_39_wide <- full_join(Ret_task_df_join, val_df) # Final data table with all computed results

datatable(group_39_wide)

```

#### Descriptives 

```{r descriptives, message=FALSE, warning=FALSE}

my_fun <- function(x, num_var){
  num_var <- enquo(num_var)

  x %>%
    summarize(avg = mean(!!num_var), n = n(), 
              sd = sd(!!num_var), se = sd/sqrt(n))
}

mean_response_time <- group_39_wide %>%
  group_by(Sex) %>%
  my_fun(MeanRT)

neg_similar_hits <- group_39_wide %>%
  group_by(Sex) %>%
  na.omit() %>% 
  my_fun(Neg_Sim_Hit)

kable(mean_response_time) %>% kable_styling()
kable(neg_similar_hits) %>% kable_styling()

```


#### Emotion Scores 

```{r emotion, message=FALSE, warning=FALSE, fig.height=4, fig.width=9}

# Gather variables and change to wide format
ret_long <- group_39_wide %>% 
  gather(key, value, Valence, Arousal, Dominance) %>%
  select(Subject, key, value, Sex) %>% 
  rename("Emotion" = key) %>% 
  arrange(desc(Subject))

ggplot(data = ret_long) +
  aes(x = value) +
  geom_histogram(bins = 30, fill = "#1f78b4") +
  labs(title = "Distribution of Emotion Response Scores",
    x = "Mean Scores",
    y = "Ratings",
    subtitle = "IAPS Scores") +
  theme_cowplot() +
  facet_wrap(vars(Emotion))
```

#### Emotion Ratings from Negative Stimuli

```{R, fig.height=4, fig.width=9}
# Negative Emotion Scores
val_task %>% 
  filter(condition == "negative") %>% 
  na.omit() %>% 
  group_by(Subject, Sex) %>% 
  summarise(Arousal = mean(arousal.RESP)
            , Dominance = mean(dominance.RESP)
            , Valence = mean(valence.RESP)) %>% 
  gather(key, value, Arousal, Dominance, Valence) %>% 
  rename("Emotion" = key) %>% 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 20, fill = "#1f78b4") +
  labs(title = "Emotion Scores from Negative Stimuli",
       y = "Response",
    subtitle = "Arousal, Dominance and Valence") +
  theme_cowplot() +
  facet_wrap(vars(Emotion))
```

As expected, negative stimuli have a higher arousal rating and lower valence rating. 


####  Emotion Ratings from Neutral Stimuli

```{r, fig.height=4, fig.width=9}
# Neutral Emotion Scores
val_task %>% 
  filter(condition == "neutral") %>% 
  na.omit() %>% 
  group_by(Subject, Sex) %>% 
  summarise(Arousal = mean(arousal.RESP)
            , Dominance = mean(dominance.RESP)
            , Valence = mean(valence.RESP)) %>% 
  gather(key, value, Arousal, Dominance, Valence) %>% 
  rename("Emotion" = key) %>% 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 20, fill = "#1f78b4") +
  labs(title = "Emotion Scores from the Neutral Condition",
    subtitle = "Arousal, Dominance and Valence") +
  theme_cowplot() +
  facet_wrap(vars(Emotion))

```

This is what we would expect. Neutral stimuli should have low arousal ratings and higher valence ratings. 

#### Histogram of Total Negative and Neutral Condition hits

```{r, fig.height=4, fig.width=9}
group_39_wide %>% 
  gather(key, value, Neutral_hits, Negative_hits) %>% # change to long format
  select(Subject, Sex, key, value) %>%
  na.omit() %>% 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 22, fill = "#1f78b4") +
  theme_cowplot() +
  facet_wrap(vars(key)) +
  geom_vline(aes(xintercept = mean(value)), linetype = "dashed") 

```


#### Negative Similar Hits (Male vs Females)

This is our overall indicator of if there is evidence of pattern separation performance. The ability to detect 
a image that was seen in the encoding task and recall if it they had seen it previously. 

```{r, fig.height=4, fig.width=9}
group_39_wide %>% 
  select(Sex, Neg_Sim_Hit, Subject) %>%
  na.omit() %>% 
  ggplot(aes(x = Neg_Sim_Hit)) +
  geom_histogram(bins = 22, fill = "#1f78b4") +
  labs(title = "Negative Similar Hits",
    subtitle = "Male and Females") +
  theme_cowplot() +
  facet_wrap(vars(Sex)) +
  geom_vline(data = na.omit(group_39_wide), aes(xintercept = mean(Neg_Sim_Hit)), linetype = "dashed")
```

#### Neutral Simiar Hits (Male vs Females)

```{r, fig.height=4, fig.width=9}

group_39_wide %>% 
  select(Sex, Neut_Sim_Hit, Subject) %>%
  na.omit() %>% 
  ggplot(aes(x = Neut_Sim_Hit)) +
  geom_histogram(bins = 22, fill = "#1f78b4") +
  labs(title = "Neutral Similar Hits",
    subtitle = "Male and Females") +
  theme_cowplot() +
  facet_wrap(vars(Sex)) +
  geom_vline(data = na.omit(group_39_wide), aes(xintercept = mean(Neut_Sim_Hit)), linetype = "dashed")
```

#### Validated vs Fake Comparisons

We need to check if the altered images were eliciting a similar emotional response as to the 
original IAPS validated images. There should be no difference between dimensions specifically, arousal and valence. If there is a difference then we dont know if we are actually eliciting the same response from participants as we want to. 

I had already processed this file previous created a file with two groups. Group "a" is the original IAPS images with the means and sds taken from the manual. Group "b" is the altered images used in the retrieval task. 

```{r message=FALSE, warning=FALSE}

IAPS_Fake_df <- read_csv("data/validation_task_df.csv", # import data
    col_types = cols(Group = col_factor(levels = c("b", 
        "a")), aromn = col_number(), dommn = col_number(), 
        valmn = col_number()), na = "0")


# Change to long format

IAPS_Fake_df <- IAPS_Fake_df %>% 
  gather(key, value, aromn, valmn, dommn) %>% 
  rename("Emotion" = key)

IAPS_Fake_df$Emotion <- as.factor(IAPS_Fake_df$Emotion)
IAPS_Fake_df$Group <- as.factor(IAPS_Fake_df$Group)

library(cowplot)

IAPS_Fake_df %>% 
  na.omit() %>% 
ggplot() +
 aes(x = Emotion, y = value, fill = Group, colour = Group) +
 geom_boxplot() +
 scale_fill_hue() +
 scale_color_hue() +
 labs(x = "Emotion", y = "Mean Arousal Score", title = "Validated vs Fake Comparison", subtitle = "Original IAPS Images with Altered Images") +
 theme_cowplot() +
 theme(legend.position = "none") +
 facet_wrap(vars(Group))

```

Boxplot seems to indicate we are getting pretty similar responses to the IAPS images. 

```{r message=FALSE, warning=FALSE}
library(ggpubr)

comparison <- compare_means(value ~ Emotion, group.by = "Group",  data = IAPS_Fake_df)

comparison


```


#### Data Analysis


1. Participants will show increased pattern separation scores on high arousal trials versus low-arousal trials.

2. We expect an interaction effect between sex and emotion, females will have a higher pattern separation score than males for negative stimuli only. 

Next post will go through how to conduct a split-plot anova and analyse the results. 